---
output: html_document
---
<center>POLS 7012 Midterm</center>
<center>Elise Blasingame</center>
<center>October 28, 2020</center>





### Problem 1

$$U(x)=\sqrt{1-x}+\sqrt{\frac{x}{W}}$$
First, we need to take the derivative of this function.

$$\frac{d}{dx}U(x)=\frac{d}{dx}(\sqrt{1-x})+\frac{d}{dx}(\sqrt{\frac{x}{W}})$$
$$\frac{d}{dx}U(x)=\frac{d}{dx}[(1-x)^\frac{1}{2}]+\frac{d}{dx}[({\frac{x}{W}})^\frac{1}{2}]$$
We can multiply each function's derivative by the derivative of the value inside the parenthesis:
$$\frac{d}{dx}U(x)=(-1)[\frac{1}{2}(1-x)^\frac{-1}{2}]+\frac{1}{2W}(\frac{x}{W})^\frac{-1}{2}$$

Now, we need to set the derivative equal to zero, and solve for x.
$$0=(-1)[\frac{1}{2}(1-x)^\frac{-1}{2}]+\frac{1}{2W}(\frac{x}{W})^\frac{-1}{2}$$
$$0=\frac{1}{\sqrt{1-x}}+\frac{1}{\sqrt\frac xW}$$
$${\sqrt{1-x}}=W{\sqrt\frac xW}$$
$${1-x}=Wx$$
$$x=\frac{1}{1+W}$$
This is the fraction that represents the tax revenue he should devote to private goods to maximize his winning coalition.

Based on this, we can say that as the winning coalition grows, the amount of revenue he should devote to private goods decreases.

### Problem 2
```{r load tidyverse}
library(tidyverse)
```
##### Step 1: Load the FIW data into R.
```{r problem 2.1}

FIW <- read.csv(file = 'data/2020_All_Data_FIW_2013-2020.csv', skip = 1)
head(FIW)
```

##### Step 2: Keep only the most recent year of data. Keep only the countries (not the territories).

```{r problem 2.2}

#library(dplyr)


FIW2020 <- filter(FIW, Edition == '2020') #Filter for 2020 data only


FIW2020 %>% 
  select(Country.Territory, C.T, Edition, A, B, C, D, E, F, G) %>% # select relevant variables
  filter(C.T == "c") %>%
  head(10) # output the first 10 rows
```

##### Step 3: Of the 25 questions, some are about democracy (i.e. are political leaders elected?) and others are about freedom/liberty (i.e. can people practice their religion without fear of persecution? is there a free and independent media? etc.). Create one variable called democracy_score, equal to the sum of all the scores on questions related to democracy, and another variable called liberty_score, equal to the sum of all the scores on questions related to liberty.

```{r problem 2.3}

FIW2020 <- FIW2020 %>% #create democracy_score
  mutate(democracy_score = A + B + C)

summary(FIW2020$democracy_score)

FIW2020 <- FIW2020 %>% #create liberty_score
  mutate(liberty_score = D + E + F + G)

summary(FIW2020$liberty_score)
```

##### Step 4: Visualize the relationship between liberty and democracy.

```{r problem 2.4}

ggplot(FIW2020, aes(x=democracy_score, y=liberty_score)) +
  geom_point(color = 'darkcyan', shape=1)+
  theme_classic() +
  labs(x = 'Democracy Scores',
       y = 'Liberty Scores')
```

##### Step 5: Split the countries into democratic and non-democratic (your choice where you split) and conduct a bivariate hypothesis test. Are the democratic countries more free on average than the non-democratic countries?

ENB: I am setting the democracy cut-off at countries with scores of =/>21

```{r problem 2.5}


FIW2020 <- FIW2020 %>%
  mutate(democracy_status = case_when(democracy_score < 21  ~ 'NonDemocratic', TRUE ~ 'Democratic'))


test1 <- FIW2020 %>%
  filter(democracy_status %in% c('NonDemocratic', 'Democratic')) %>%
  t.test(liberty_score ~ democracy_status, data = .)


test1
```

##### Because the p-value is very low, we can reject the null hypothesis and conclude that Democratic Countries are typically 'more free' than Non-Democratic Countries. 